---
title: 'CMTH 642 Data Analytics: Advanced Methods'
output:
  pdf_document: default
  html_document: default
  word_document: default
---
<center> <h1> Assignment 1 (10%) </h1> </center>
<center>  <h3> [Tara Shingadia] </h2> </center>
<center> <h3> [DJ0 & 500847789] </h2> </center>
---

#### 1. Read the csv files in the folder. (4 points)
```{r}
# INSERT YOUR ANSWER HERE
USDA_Macro <- read.csv("USDA_Macronutrients.csv", header=TRUE)
USDA_Micro <- read.csv("USDA_Micronutrients.csv", header=TRUE)
```

#### 2. Merge the data frames using the variable "ID". Name the Merged Data Frame "USDA". (4 points)
```{r}
# INSERT YOUR ANSWER HERE
USDA <- merge(USDA_Macro, USDA_Micro, by='ID')
```

#### 3. Check the datatypes of the attributes. Delete the commas in the Sodium and Potasium records. Assign Sodium and Potasium as numeric data types. (4 points)
```{r}
# INSERT YOUR ANSWER HERE
str(USDA)
USDA$Potassium <- as.numeric(gsub(",","",USDA$Potassium))
USDA$Sodium <- as.numeric(gsub(",","",USDA$Sodium))
```

#### 4. Remove records (rows) with missing values in more than 4 attributes (columns). How many records remain in the data frame? (4 points)
```{r}
# INSERT YOUR ANSWER HERE
missing_value <- rowSums(is.na(USDA))
missing4 <- missing_value > 4
USDA1 <- USDA[!missing4,]
nrow(USDA1)
```

#### 5. For records with missing values for Sugar, Vitamin E and Vitamin D, replace missing values with mean value for the respective variable. (4 points)
```{r}
# INSERT YOUR ANSWER HERE
USDA1$Sugar[is.na(USDA1$Sugar)] <- mean(USDA$Sugar, na.rm = TRUE)
USDA1$VitaminE[is.na(USDA1$VitaminE)] <- mean(USDA$VitaminE, na.rm = TRUE)
USDA1$VitaminD[is.na(USDA1$VitaminD)] <- mean(USDA$VitaminD, na.rm = TRUE)
USDA1
```

#### 6. With a single line of code, remove all remaining records with missing values. Name the new Data Frame "USDAclean". How many records remain in the data frame? (5 points)
```{r}
# INSERT YOUR ANSWER HERE
USDAclean <- USDA1[complete.cases(USDA1),]
nrow(USDAclean)
```

#### 7. Which food has the highest sodium level? (5 points)
```{r}
# INSERT YOUR ANSWER HERE
USDAclean$Description[which.max(USDAclean$Sodium)]
```

#### 8. Create a histogram of Vitamin C distribution in foods. (5 points)
```{r}
# INSERT YOUR ANSWER HERE
Vitamin_C <- USDAclean$VitaminC
hist(Vitamin_C, xlim = c(0,120), breaks = 150, ylim = c(0,6000), col = 7, main = "Vitamin C Distribution in Food", xlab = "Vitamin C (mg)")
```

#### 9. Create one boxplot to illustrate the distribution of values for TotalFat, Protein and Carbohydrate. (5 points)
```{r}
# INSERT YOUR ANSWER HERE
colours = c('firebrick1','darkorange','goldenrod1')
boxplot(USDAclean$TotalFat, USDAclean$Protein, USDAclean$Carbohydrate, names = c("Total Fat","Protein", "Carbohydrates"), col = colours)
```

#### 10. Create a scatterplot to illustrate the relationship between a food's TotalFat content and its Calorie content. (5 points)
```{r}
# INSERT YOUR ANSWER HERE
library(ggplot2)
ggplot(data = USDAclean, aes(x = TotalFat, y = Calories)) + geom_point(color = "deepskyblue2")
```

#### 11. Add a variable to the data frame that takes value 1 if the food has higher sodium than average, 0 otherwise. Call this variable HighSodium. Do the same for High Calories, High Protein, High Sugar, and High Fat. How many foods have both high sodium and high fat? (5 points)
```{r}
# INSERT YOUR ANSWER HERE
USDAclean$HighSodium <- ifelse(USDAclean$Sodium > mean(USDAclean$Sodium), 1, 0)
USDAclean$HighCalories <- ifelse(USDAclean$Calories > mean(USDAclean$Calories), 1, 0)
USDAclean$HighProtein <- ifelse(USDAclean$Protein > mean(USDAclean$Protein), 1, 0)
USDAclean$HighSugar <- ifelse(USDAclean$Sugar > mean(USDAclean$Sugar), 1, 0)
USDAclean$HighFat <- ifelse(USDAclean$TotalFat > mean(USDAclean$TotalFat), 1, 0)

length(which(USDAclean$HighSodium == 1 & USDAclean$HighFat == 1))
```

#### 12. Calculate the average amount of iron, for high and low protein foods. (5 points)
```{r}
# INSERT YOUR ANSWER HERE
mean(USDAclean[USDAclean$HighProtein == 1, 'Iron'])
mean(USDAclean[USDAclean$HighProtein == 0, 'Iron'])
```

#### 13. Create a function for a "HealthCheck" program to detect unhealthy foods. Use the algorithm flowchart below as a basis. (5 points)
```{r}
require(jpeg)
img<-readJPEG("HealthCheck.jpg")
plot(1:4, ty = 'n', ann = F, xaxt = 'n', yaxt = 'n')
rasterImage(img,1,1,4,4)
```
```{r}
# INSERT YOUR ANSWER HERE
HealthCheck <- function(x,y,z){ifelse(x == 1, ifelse(y == 1, ifelse(z == 1, "Fail", "Pass"), "Pass"), "Pass")}
```

#### 14. Add a new variable called HealthCheck to the data frame using the output of the function. (5 points)
```{r}
# INSERT YOUR ANSWER HERE
USDAclean$HealthCheck <- HealthCheck(USDAclean$HighSodium, USDAclean$HighSugar, USDAclean$HighFat)
```

#### 15. How many foods in the USDAclean data frame fail the HealthCheck? (5 points)
```{r}
# INSERT YOUR ANSWER HERE
nrow(USDAclean[USDAclean$HealthCheck == "Fail",])
```

#### 16. Visualize the correlation among Calories, Protein, Total Fat, Carbohydrate, Sodium and Cholesterol. (5 points)
```{r}
# INSERT YOUR CODE HERE.
library(corrplot)
Attributes <- USDAclean[c("Calories", "Protein", "TotalFat", "Carbohydrate", "Sodium", "Cholesterol")]
H = cor(Attributes)
corrplot(H, method = 'number')
```

#### 17. Is the correlation between Calories and Total Fat statistically significant? Why? (5 points)
```{r}
# INSERT YOUR CODE HERE AND EXPLAIN YOUR ANSWER AS A COMMENT.
test <- cor.test(USDAclean$Calories, USDAclean$TotalFat)
test
#The correlation between calories and total fat is statistically significant because it has a p value of 2.2e-16 which is < 0.05 at a 95% confidence interval
```

#### 18. Create a Linear Regression Model, using Calories as the dependent variable Protein, Total Fat, Carbohydrate, Sodium and Cholesterol as the independent variables. (4 points)
```{r}
# WRITE YOUR ANSWER AS A COMMENT.
model1 <- lm(Calories ~ Protein + TotalFat + Carbohydrate + Sodium + Cholesterol, data = USDAclean)
summary(model1)
```

#### 19. Which independent variable is the least significant? Why? (4 points)
```{r}
# INSERT YOUR CODE HERE AND EXPLAIN YOUR ANSWER AS A COMMENT.
#Sodium is the least significant independent variable because according to the significance codes in the summary table above, it falls between the range of [0.1, 1] which is not significant.
```

#### 20. Create a new model by using only the significant independent variables. (4 points)
```{r}
# INSERT YOUR CODE HERE.
model2 <- lm(Calories ~ Protein + TotalFat + Carbohydrate + Cholesterol, data = USDAclean)
summary(model2)
```

#### 21. A new product is just produced with the following data: Protein=0.1, TotalFat=37, Carbohydrate=400, Cholesterol=75, Sugar=NA, Calcium=35, Iron=NA, Potassium=35, VitaminC=10,	VitaminE=NA, VitaminD=NA. Based on the new model you created, what is the predicted value for Calories? (4 points)
```{r}
# INSERT YOUR CODE HERE.
new_product <- data.frame()
predict(model2, )
```

#### 22. If the Carbohydrate amount increases from 400 to 40000 (10000% increase), how much change will occur on Calories in percent? Explain why? (4 points)
```{r}
# INSERT YOUR CODE HERE AND EXPLAIN YOUR ANSWER AS A COMMENT.
```


This is the end of Assignment 1

Ceni Babaoglu, PhD
